---
editor_options:
  markdown:
    wrap: 72
---

# Probability and Statistics

# Lab Assignment 1: Naive Bayes Classifier

### *Andrii Yaroshevych, Dmytro Vasylkiv, Vitalii Petrychko*

## Introduction

During the past three weeks, you learned a couple of essential notions
and theorems, and one of the most important among them is the *Bayes
theorem*.

One of its applications is **Naive Bayes classifier**, which is a
probabilistic classifier whose aim is to determine which class some
observation probably belongs to by using the Bayes formula:
$$\mathsf{P}(\mathrm{class}\mid \mathrm{observation})=\frac{\mathsf{P}(\mathrm{observation}\mid\mathrm{class})\mathsf{P}(\mathrm{class})}{\mathsf{P}(\mathrm{observation})}$$

Under the strong independence assumption, one can calculate
$\mathsf{P}(\mathrm{observation} \mid \mathrm{class})$ as
$$\mathsf{P}(\mathrm{observation}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i), \qquad \mathsf{P}(\mathrm{observation} \mid \mathrm{class}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i \mid \mathrm{class}),$$
where $n$ is the total number of features describing a given
observation. Thus, $\mathsf{P}(\mathrm{class}|\mathrm{observation})$ now
can be calculated as

$$\mathsf{P}(\mathrm{class} \mid \mathrm{\mathrm{observation}}) = \mathsf{P}(\mathrm{class})\times \prod_{i=1}^{n}\frac{\mathsf{P}(\mathrm{feature}_i\mid \mathrm{class})}{\mathsf{P}(\mathrm{feature}_i)}\tag{1}$$

All the terms on the right-hand side can be estimated from the data as
respective relative frequencies;\
see [this
site](https://monkeylearn.com/blog/practical-explanation-naive-bayes-classifier/)
for more detailed explanations.

## Data description

There are 5 datasets uploaded on the cms.

To determine your variant, take your team number from the list of teams
on cms and take *mod 5* - this is the number of your data set.

-   **0 - authors** This data set consists of citations of three famous
    writers: Edgar Alan Poe, Mary Wollstonecraft Shelley and HP
    Lovecraft. The task with this data set is to classify a piece of
    text with the author who was more likely to write it.

-   **1 - discrimination** This data set consists of tweets that have
    discriminatory (sexism or racism) messages or of tweets that are of
    neutral mood. The task is to determine whether a given tweet has
    discriminatory mood or does not.

-   **2 - fake news** This data set contains data of American news: a
    headline and an abstract of the article. Each piece of news is
    classified as fake or credible. The task is to classify the news
    from test.csv as credible or fake.

-   **3 - sentiment** All the text messages contained in this data set
    are labeled with three sentiments: positive, neutral or negative.
    The task is to classify some text message as the one of positive
    mood, negative or neutral.

-   **4 - spam** This last data set contains SMS messages classified as
    spam or non-spam (ham in the data set). The task is to determine
    whether a given message is spam or non-spam.

Each data set consists of two files: *train.csv* and *test.csv*. The
first one you will need find the probabilities distributions for each of
the features, while the second one is needed for checking how well your
classifier works.

```{r}
# here goes a list of recommended libraries,
# though you may install other ones if they are needed
#install.packages("tidytext")
#install.packages("readr")
#install.packages("dplyr")
#install.packages("ggplot2")
library(tidytext)
library(readr)
library(dplyr)
library(ggplot2)
```

## Instructions

-   The first step is data pre-processing, which includes removing
    punctuation marks and stop words

-   represent each message as a bag-of-words

-   using the training set, calculate all the conditional probabilities
    in formula (1)

-   use those to predict classes for messages in the test set

-   evaluate effectiveness of the classifier by calculating the
    corresponding metrics

-   shortly summarize your work

-   do not forget to submit both the (compiled) Rmd source file and the
    .html output

### Data pre-processing

-   Read the *.csv* data files.
-   Сlear your data from punctuation or other unneeded symbols.
-   Clear you data from stop words. You don't want words as is, and, or
    etc. to affect your probabilities distributions, so it is a wise
    decision to get rid of them. Find list of stop words in the cms
    under the lab task.
-   Represent each test message as its bag-of-words. Here:
    <https://machinelearningmastery.com/gentle-introduction-bag-words-model/>
    you can find general introduction to the bag-of-words model and
    examples on to create it.
-   It is highly recommended to get familiar with R dataframes, it would
    make the work much easier to do.
-   Useful links:
    -   <https://steviep42.github.io/webscraping/book/bagofwords.html#tidytext> -
        example of using *tidytext* to count frequencies of the words.
    -   Basics of Text Mining in R:
        <http://rstudio-pubs-static.s3.amazonaws.com/256588_57b585da6c054349825cba46685d8464.html>
        . Note that it also includes an example on how to create a bag
        of words from your text document.

```{r}
list.files(getwd())
list.files("data/2-fake_news")
```

```{r}
test_path <- "data/2-fake_news/test.csv"
train_path <- "data/2-fake_news/train.csv"

stop_words_file <- read_file("stop_words.txt")
# https://stackoverflow.com/questions/27195912/why-does-strsplit-return-a-list
stop_words <- strsplit(stop_words_file, split='\n')
stop_words <- stop_words[[1]]
View(stop_words)
```

```{r}
train <- read.csv(file = train_path, stringsAsFactors = FALSE)
test <- read.csv(file = test_path, stringsAsFactors = FALSE)
```

```{r}
tidy_text <- unnest_tokens(train, 'splitted', 'Body', token="words", to_lower = TRUE) %>%
             filter(!splitted %in% stop_words)

tidy_text %>% count(splitted, sort=TRUE)
```

### Data visualization

```{r}
fake_news <- tidy_text %>%
             filter(Label == "fake") %>%
             count(splitted, sort=TRUE) %>%
             top_n(15)

credible_news <- tidy_text %>%
                 filter(Label == "credible") %>%
                 count(splitted, sort=TRUE) %>%
                 top_n(15)

ggplot(fake_news, aes(x = n, y = reorder(splitted, n))) +
    geom_col() +
    labs(x = "Frequency", y = "Word", title = "Top 15 words in fake news") +
    theme_minimal()

ggplot(credible_news, aes(x = n, y = reorder(splitted, n))) +
    geom_col() +
    labs(x = "Frequency", y = "Word", title = "Top 15 words in credible news") +
    theme_minimal()
```

## Classifier implementation

```{r}
naiveBayes <- setRefClass("naiveBayes",

       fields = list(
         data = "data.frame",
         fake_data = "data.frame",
         credible_data = "data.frame"
       ),

       methods = list(
                    fit = function(data)
                    {
                        data <<- data
                        fake_data <<- data %>% filter(Label=="fake") %>% count(splitted, sort=TRUE)
                        credible_data <<- data %>% filter(Label=="credible") %>% count(splitted, sort=TRUE)

                        fake_data <<- fake_data
                        credible_data <<- credible_data
                    },
                    
                    predict = function(message)
                    {
                      
                        number_of_fake_news <- nrow(data %>% filter(Label=="fake"))
                        number_of_credible_news <- nrow(data %>% filter(Label=="credible"))

                        fake_news_prob <- number_of_fake_news/(number_of_fake_news + number_of_credible_news)
                        credible_news_prob <- number_of_credible_news/(number_of_fake_news + number_of_credible_news)
                        #print(message)
                        message <- tolower(message)
                        #print(message)
                        entries <- strsplit(message, split=" ")[[1]]
                        entries <- entries[!(entries %in% stop_words)]

                        #fake_conditional_prob <- 1
                        #credible_conditional_prob <- 1

                        fake_news_word_count <- sum(fake_data$n)
                        credible_news_word_count <- sum(credible_data$n)
                        result<-1

                        total_word_count <- nrow(data)

                        for (word in entries) {
                            number_of_entries_fake <- ifelse(
                              word %in% fake_data$splitted,
                              fake_data %>% filter(splitted==word) %>% pull(n), 0)
                            fake_word_conditional_prob <- (number_of_entries_fake + 1)/
                              (fake_news_word_count + total_word_count)
                            number_of_entries_credible <- ifelse(
                              word %in% credible_data$splitted,
                              credible_data %>% filter(splitted==word) %>% pull(n), 0)

                           credible_word_conditional_prob <- (number_of_entries_credible + 1)/
                             (credible_news_word_count + total_word_count)
                           result<-result*(fake_word_conditional_prob/credible_word_conditional_prob)

                        }
                        result<-result*(fake_news_prob/credible_news_prob)


                        if (result > 1) {
                            return("fake")
                        } else {
                            return("credible")
                        }
                    },

                    score = function(test)
                    {
                         right_predictions=0
                         total_titles_count=nrow(test)
                         for (row_index in 1:total_titles_count){
                           message <- test[row_index, "Body"]
                           result <- test[row_index, "Label"]
                           if (predict(message)==result){
                             right_predictions=right_predictions+1
                          }
                         }
                         #result <- data.frame(test)
                         #result$prediction <- lapply(result$Body,MARGIN=1, predict)
                         #View(result)
                         #}
                         #test$prediction <- lapply(X=test[,c('Body')], MARGIN=1, FUN=self$predict)
                         #View(test)
                         
                         #correct_values <- nrow(test.which(test$Label==test$prediction))
                         #return(correct_values/nrow(test))
                         return(right_predictions/total_titles_count)
                    }
))

model <- naiveBayes()
model$fit(tidy_text)
#prediction <- model$predict("Will")
#print(prediction)
#result <- data.frame(test)
#result$prediction <- apply(result["Body"],MARGIN=1, model$predict)
#View(result)
                         #}
print(model$score(test))
```

\`\`\`

## Measure effectiveness of your classifier

-   Note that accuracy is not always a good metric for your classifier.
    Look at precision and recall curves, F1 score metric.
-   Visualize them.
-   Show failure cases.

## Conclusions

Summarize your work by explaining in a few sentences the points listed
below.

-   Describe the method implemented in general. Show what are
    mathematical foundations you are basing your solution on.
-   List pros and cons of the method. This should include the
    limitations of your method, all the assumption you make about the
    nature of your data etc.
